Avoiding accuracy-limiting pitfalls in the study of protein-ligand interactions with isothermal titration calorimetry

### 2016-01-03 manuscript/itc-worksheet.tex
Update funding acknowledgments.
Incorporate edits to backlash effect passage from Joel

{\bf Syringe backlash and the first-injection anomaly.}
Our VP-ITC instrument has a syringe assembly that utilizes a worm gear which, after the recommended purge-refill process, will cause a titrant shortfall in the first injection unless a ``down syringe'' command is issued prior to loading the syringe into the sample cell~\cite{tellinghuisen:anal-biochem:2004:first-injection-anomaly}; we therefore executed a 10~$\mu$L ``down syringe'' command immediately after the purge-refill cycle.
\color{red}
Note that newer calorimeters such as the Malvern/GE/MicroCal iTC-200 that utilize a $\sim$200 $\mu$L cell volume have also been observed to suffer from this syringe backlash phenomena;
a recent analysis found a 1.3$\mu$L shortfall in titrant delivery contributing as much as 8\% to error in $K$ and $n$~\cite{tellinghuisen:2015:bba-general:optimizing-itc}.
Because the stoichiometric endpoint is reached after a much smaller total amount of titrant has been delivered for these low-volume instruments---the iTC-200 syringe has a total usable volume of only 35 $\mu$L---the importance of ensuring a ``down syringe'' command is issued prior to loading the syringe into the sample cell is critical for achieving high accuracy.
\color{black}

@@

JDC and SEB acknowledge funding from the Memorial Sloan Kettering Cancer Center, as well as from the California Institute for Quantitative Biosciences (QB3) at the University of California during earlier stages of this work.
The authors are grateful to Brian K.~Shoichet (UCSF) for the use of his laboratory facilities for this study, as well as Allison Doak (UCSF) and Matthew Merski (UCSF) for their assistance.
The authors thank Patrick Grinaway, Daniel Parton, and Ari\"{e}n Sebastian Rustenburg (MSKCC) for stimulating the publication of this work through their subsequent work on this system, and Antonio Luz and Fraser Glickman (Rockefeller High-Throughput and Spectroscopy Resource Center) for their assistance and expertise with automated calorimetry.

>>>>

{\bf Syringe backlash and the first-injection anomaly.}
Our VP-ITC instrument has a syringe assembly that utilizes a worm gear which, after the recommended purge-refill process, will cause a titrant shortfall in the first injection (due to ``syringe backlash'') unless a ``down syringe'' command is issued prior to loading the syringe into the sample cell~\cite{tellinghuisen:anal-biochem:2004:first-injection-anomaly}; we therefore executed a 10~$\mu$L ``down syringe'' command immediately after the purge-refill cycle.
This procedure may be even more important for newer, small-volume instruments.  
A recent analysis of data collected on a  Malvern/GE/MicroCal iTC-200 (cell volume $V_0 \approx 200$ $\mu$L) found a 1.3-$\mu$L backlash error, neglect of which amounts to an 8\% relative error in $K$ and $n$~\cite{tellinghuisen:2015:bba-general:optimizing-itc}.  
This is much greater than the backlash-induced relative error in $K$ and $n$ for the VP-ITC because typical stoichiometric endpoint volumes are much smaller on the small-volume instruments (the total usable titrant volume being just 35 $\mu$L for the iTC-200).

@@

The authors are grateful to Brian K.~Shoichet (UCSF) for the use of his laboratory facilities for this study, as well as Allison Doak (UCSF) and Matthew Merski (UCSF) for their assistance.
The authors thank Patrick Grinaway, Daniel Parton, and Ari\"{e}n Sebastian Rustenburg (MSKCC) for stimulating the publication of this work through their subsequent work on this system, and Antonio Luz and Fraser Glickman (Rockefeller High-Throughput and Spectroscopy Resource Center) for their assistance and expertise with automated calorimetry.
JDC and SEB acknowledge funding from the Memorial Sloan Kettering Cancer Center, as well as from the California Institute for Quantitative Biosciences (QB3) at the University of California during earlier stages of this work.
JDC also acknowledges a Louis V.~Gerstner Young Investigator Award and NIH core grant P30-CA008748 for partial funding during the course of this work.
